Rectangulotopes

### 2024-04-22 draft/rectangulotopes.tex
l23

\begin{lemma}
  Let $R$ be a rectangulation of size~$n$, let $S$ and $T$ be its source and target trees, and let~$([n],\prec_s)$ be the associated strong order.
  Given a pair $i,j\in [n]$ such that $i<j$ and $i\prec_s j$, we have
  \[
    cv^{T,S}_{i,j} > 0 \quad \Longrightarrow \quad i \, \backslash \, j.
  \]
\end{lemma}

>>>>

\begin{lemma}
  Let $R$ be a rectangulation of size~$n$, let $S$ and $T$ be its source and target trees, and let~$([n],\prec_s)$ be the associated strong order.
  Given a pair $i,j\in [n]$ such that $i<j$ and $i\prec_s j$, we have
  \[
    cv^{T,S}_{i,j} > 0 \quad \Longrightarrow \quad i \, \backslash \, j.
  \]
\end{lemma}
\begin{proof}
  Consider the two rectangles $r_i$ and $r_j$ with inorder labels $i$ and $j$ in $R$.
  We first observe that the source vertex $s(r_j)$ is not contained in the interior of the lower left quadrant of origin $t(r_i)$.
  Indeed, if we suppose otherwise, then the line segment from $s(r_j)$ to $t(r_i)$ has positive slope, which implies that $i\tri j$, which in turn implies $i\succ_s j$, contradicting our hypothesis.
  We now proceed to a simple case analysis.

  First suppose that $r_i$ and $r_j$ are disjoint, and that none of their edges lie on the same segment of $R$.
  Then we claim that $cv^{T,S}_{i,j} = 0$.
  Otherwise, suppose that $r_i$ and $r_j$ are disjoint, but they share a segment.
  Without loss of generality (the other case is symmetric), let us suppose that they share a vertical segment.

  Hence it cannot be the case that $r_i$ and $r_j$ are disjoint, and they must be touching.
  Then, from our first observation, it can only be the case that $r_i$ is left of $r_j$, or below $r_j$, hence that $i\, \backslash \, j$, as claimed.
  \end{proof}
